Sequence Specific Modeling of Cell-Free Protein Synthesis

### 2017-03-06 Paper_v1.tex
MM updates

\subsection{Model Derivation}
The goal of this work was first to construct a modeling framework to describe cell-free protein synthesis systems and to examine its performance in productivity, energy efficiency and carbon yield for a protein of interest.
One mathematical framework that has found wide use in modeling metabolism is constraints based models such as flux balance analysis (FBA).
FBA can predict how cells utilize nutrients to produce products by using the biochemical stoichiometry and thermodynamical feasibility under pseudo steady-state conditions.
Traditionally, FBA is used to model \textit{in~vivo} processes; however, cell-free systems do not have growth associated reactions or transport through the cell membrane.
Thus, to model cell-free metabolism we constructed a stoichiometric network by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa}, and incorporating amino acid synthesis and transcription/translation associated reactions \cite{Allen:2003aa} for a protein of interest to be expressed.
The network consisted of 272 reactions and 145 species.
The core metabolism is shown in Fig.~\ref{fig:network}A.
We developed a cell-free sequence specific flux balance analysis (ssFBA) with a detailed promoter model \cite{Moon:2012aa} to examine the performance of CFPS.
We first validated the ssFBA approach by comparing simulated and measured concentrations of two proteins from two different cell-free \textit{E.~coli} extract systems.
The first protein, chloramphenicol acetyltransferase (CAT), was produced under a T7 promoter in a glucose/NMP cell-free system \cite{2005_calhoun_BiotechnologyProgress} for a duration of 1 hour under glucose consumption (Fig.~\ref{fig:network}B).
The second protein, dual emission green fluorescent protein (deGFP), was produced under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for a duration of 8 hours under maltose consumption (Fig.~\ref{fig:network}C).
The ssFBA simulations predicted the production of both proteins for the duration of both CFPS batch reactions.
Uncertainty in experimental factors such as RNA polymerase, ribosome concentrations, elongation rates and the upper bounds for oxygen and glucose consumption rates did not alter the qualitative performance of the model.
Thus, the metabolic network and molecular description of trancription and translation were consistent with experimental measurements.
Next, ssFBA predicted deGFP production as a function of plasmid concentration (Fig.~\ref{fig:network}D).
Concentration of deGFP at each plasmid concentration was calculated by multiplying the flux of deGFP synthesis by the active time of production, approximately 8 hours in TXTL 2.0 \cite{Garamella:2016aa}.
The mean of the ensemble shows a good prediction against the measured deGFP levels, even though it under predicted deGFP concentration at the saturating point of 5 nM of plasmid concentration.
However, the ensemble and the mean of the ensemble captured the overall saturating dynamics of deGFP production as a function of plasmid concentration.
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/ssFBA_network.pdf}
\caption{Sequence specific flux balance analysis. A. Core metabolic network with glycolysis, pentose phoshate pathway, TCA cycle, Entner-Doudoroff pathway. Thick gray arrows indicate withdrawal of precursors for amino acid synthesis. B. CAT production under a T7 promoter in CFPS \textit{E.~coli} extract for 1 h under glucose consumption. C. deGFP production under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for 8 h under maltose consumption. D. Predicted deGFP concentration at different plasmid concentrations versus measurements of deGFP synthesized in TXTL 2.0. 95\% CI (blue region) over the ensemble of 100 sets, mean of the ensemble (black line), and experimental measurements (dots).}
\label{fig:network}
\end{figure}

@@

>>>>

\subsection{Model Derivation}
The goal of this work was first to construct a modeling framework to describe cell-free protein synthesis systems and to examine its performance in productivity, energy efficiency and carbon yield for a protein of interest.
One mathematical framework that has found wide use in modeling metabolism is constraints based models such as flux balance analysis (FBA).
FBA can predict how cells utilize nutrients to produce products by using the biochemical stoichiometry and thermodynamical feasibility under pseudo steady-state conditions.
Traditionally, FBA is used to model \textit{in~vivo} processes; however, cell-free systems do not have growth associated reactions or transport through the cell membrane.
Thus, to model cell-free metabolism we constructed a stoichiometric network by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa}, and incorporating amino acid synthesis and transcription/translation associated reactions \cite{Allen:2003aa} for a protein of interest to be expressed.
The network consisted of 272 reactions and 145 species.
The core metabolism is shown in Fig.~\ref{fig:network}A.
We developed a cell-free sequence specific flux balance analysis (ssFBA) with a detailed promoter model \cite{Moon:2012aa} to examine the performance of CFPS.
We first validated the ssFBA approach by comparing simulated and measured concentrations of two proteins from two different cell-free \textit{E.~coli} extract systems.
The first protein, chloramphenicol acetyltransferase (CAT), was produced under a T7 promoter in a glucose/NMP cell-free system \cite{2005_calhoun_BiotechnologyProgress} for a duration of 1 hour under glucose consumption (Fig.~\ref{fig:network}B).
The second protein, dual emission green fluorescent protein (deGFP), was produced under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for a duration of 8 hours under maltose consumption (Fig.~\ref{fig:network}C).
The ssFBA simulations predicted the production of both proteins for the duration of both CFPS batch reactions.
Uncertainty in experimental factors such as RNA polymerase, ribosome concentrations, elongation rates and the upper bounds for oxygen and glucose consumption rates did not alter the qualitative performance of the model.
Thus, the metabolic network and molecular description of trancription and translation were consistent with experimental measurements.
Next, ssFBA predicted deGFP production as a function of plasmid concentration (Fig.~\ref{fig:network}D).
Concentration of deGFP at each plasmid concentration was calculated by multiplying the flux of deGFP synthesis by the active time of production, approximately 8 hours in TXTL 2.0 \cite{Garamella:2016aa}.
The mean of the ensemble shows a good prediction against the measured deGFP levels, even though it under predicted deGFP concentration at the saturating point of 5 nM of plasmid concentration.
However, the ensemble and the mean of the ensemble captured the overall saturating dynamics of deGFP production as a function of plasmid concentration.
\includegraphics[width=1.00\textwidth]{./Figures/ssFBA_network.pdf}
\caption{Sequence specific flux balance analysis. A. Core metabolic network with glycolysis, pentose phoshate pathway, TCA cycle, Entner-Doudoroff pathway. Thick gray arrows indicate withdrawal of precursors for amino acid synthesis. B. CAT production under a T7 promoter in CFPS \textit{E.~coli} extract for 1 h under glucose consumption. Error bars denote the standard deviation of experimental measurements. C. deGFP production under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for 8 h under maltose consumption. Error bars denote a 10\% coefficient of variation. D. Predicted deGFP concentration at different plasmid concentrations versus measurements of deGFP synthesized in TXTL 2.0. 95\% CI (blue region) over the ensemble of 100 sets, mean of the ensemble (black line), and experimental measurements (dots).}
\label{fig:network}

@@

\subsection*{Cell-free protein synthesis.}
The glucose/NMP cell-free protein synthesis reactions were performed at a volume of 15 $\mu$L in 1.5-mL eppendorf tubes and incubated in a humidified incubator. 
Plasmid pK7CAT was used as the DNA template for chloramphenical acetyl transferase (CAT) expression by placing the \emph{cat} gene between the T7 promoter and the T7 terminator \cite{Kigawa1995}.
The plasmid was isolated and purified using a Plasmid Maxi Kit (Qiagen, Valencia CA). 
Cell-free protein synthesis of CAT protein was performed at 37 $^{\circ}$C for 3 h.

S30 extract was prepared from E. coli strain KC6 (A19 $\delta$tonA $\delta$tnaA $\delta$speA $\delta$endA $\delta$sdaA $\delta$sdaB $\delta$gshA met+). 
This K12-derivative has several gene deletions to stabilize amino acid concentrations during the cell-free reaction. 
The KC6 strain was grown to approximately 3 OD595 in a 10-L fermenter (B. Braun, Allentown PA) on defined media with glucose as the carbon source and with the addition of 13 amino acids (alanine, arginine, cysteine, serine, aspartate, glutamate, and glutamine were excluded) \cite{Zawada:2003}. 
Crude S30 extract was made as previously described \cite{Jewett:2002}.

The standard protein synthesis reaction was conducted according to the PANOxSP protocol with slight modifications from that described previously \cite{BIT:BIT20026}. 
Unless otherwise noted, all reagents were purchased from Sigma (St. Louis, MO). 
The initial mixture includes 1.2 mM ATP; 0.85 mM each of GTP, UTP, and CTP; 30 mM phosphoenolpyruvate (Roche, Indianapolis IN); 130 mM potassium glutamate; 10 mM ammonium glutamate; 16 mM magnesium glutamate; 50 mM HEPES-KOH buffer (pH 7.5); 1.5 mM spermidine; 1.0 mM putrescine; 34 $\mu$g/mL folinic acid; 170.6 $\mu$g/mL E. coli tRNA mixture (Roche, Indianapolis IN); 13.3 $\mu$g/mL pK7CAT plasmid; 100 $\mu$g/mL T7 RNA polymerase; 20 unlabeled amino acids at 2 mM each; 5 $\mu$M l-[U-\textsuperscript{14}C]-leucine (Amersham Pharmacia, Uppsala Sweden); 0.33 mM nicotinamide adenine dinucleotide (NAD); 0.26 mM coenzyme A (CoA); 2.7 mM sodium oxalate; and 0.24 volumes of E. coli S30 extract.
This reaction was modified for the energy source used such that glucose reactions have 30 mM glucose in place of PEP, and glutamate reactions do not have a secondary energy source added since glutamate is already present in high concentrations from the reaction salts. Sodium oxalate was not added since it has a detrimental effect on protein synthesis and ATP concentrations when using glucose or an early glycolytic intermediate as the energy source \cite{BIT:BIT1121}. The HEPES buffer (pKa $\sim$ 7.5) was replaced with Bis-Tris (pKa $\sim$ 6.5). In addition, the magnesium glutamate concentration was reduced to 8 mM for the glucose and glutamate reactions since a lower magnesium optimum is found when using a nonphosphorylated energy source \cite{BIT:BIT20026}. Finally, 10 mM phosphate was added to some reactions as described below in the form of potassium phosphate dibasic adjusted to pH 7.2 with acetic acid.

\subsection*{Measurements of protein product and metabolites.}
Cell-free reaction samples were quenched at specific timepoints with equal volumes of ice-cold 150 mM sulfuric acid to precipitate proteins. 
Protein synthesis of CAT was determined from the total amount of \textsuperscript{14}C-leucine-labeled product by trichloroacetic acid precipitation followed by scintillation counting as described previously \cite{2005_calhoun_BiotechnologyProgress}.
Samples were centrifuged for 10 min at 12,000g and 4$^{\circ}$C. 
The supernatant was collected for high performance liquid chromatography (HPLC) analysis. 
HPLC analysis (Agilent 1100 HPLC, Palo Alto CA) was used to separate nucleotides and organic acids, including glucose. Compounds were identified and quantified by comparison to known standards for retention time and UV absorbance (260 nm for nucleotides and 210 nm for organic acids) as described previously \cite{2005_calhoun_BiotechnologyProgress}.  
Calibration standards were routinely run for improved quantification of compounds. 
The standard compounds quantified with a refractive index detector included inorganic phosphate, glucose, and acetate. 
The compounds quantified with the UV detector included pyruvate, malate, succinate, and lactate.
The stability of the amino acids in the cell extract was determined using a Dionex Amino Acid Analysis (AAA) HPLC System (Sunnyvale, CA) that separates amino acids by gradient anion exchange (AminoPac PA10 column).
Compounds were identified with pulsed amperometric electrochemical detection and by comparison to known standards.
